curiobaby_drop model analysis

### 2021-01-31 analysis/model_analysis_v2.Rmd
cleaning up model evaluation report

## Load data

```{r, load-data, echo=F}

@@

mdat <- mdat %>% left_join(mdat_eng)

# selects given feature and returns wide df that can be compared with behavioral data
get_feature_df <- function(mdat, fname) {
  fdf <- mdat[,c(fname,"drop","target","relation")] %>%
    pivot_wider(id_cols = c("drop","relation"), names_from = "target", values_from = fname) %>%
    arrange(relation,drop)
}

# Fit to adult choice proportions

```{r, echo=F}
load(file=here("paper/processed_data.RData"))

@@

evalFit <- function(feat_df, hum_df, beta, cor_objective=F, relation=NA) {
  hum_df <- get_model_target_choice_props(feat_df, hum_df, beta)
  if(!is.na(relation)) {
    hum_df <- subset(hum_df, relation==relation)
  }
  r = with(hum_df, cor(chose_target, Model_targ_prop))
  mse = with(hum_df, sum((chose_target - Model_targ_prop)^2)) / nrow(hum_df) 
  if(cor_objective) mse = mse - r # also try to minimize -correlation
  return(mse) # return negative cor since DEoptim minimizes
}

require(DEoptim)

@@

```{r fit-models-to-all-trials-mse, echo=F, eval=FIT_MODELS}
#adult_fits = list()
#child_fits = list()
all_trials_mse <- tibble()

@@

## Fit separately to contain and support trials (MSE + r)

NOT DONE

@@

## Fits to all trials (MSE + r)

@@

```{r, echo=F}
fit_tab <- supp_trials_mse %>% pivot_wider( names_from = Age, values_from = c(beta, mse)) %>%
  arrange(mse_adults)

@@

@@

```{r, eval=F}
fits %>% ggplot(aes(x=beta, y=mse, color=AgeGroup)) +
  #geom_text(aes(label=feature)) + 
  geom_point(alpha=.5) +
  theme_minimal()
```
So generally, children's choices are best explained by small values of beta (.5, the minimum tested).

What do the predicted choice proportions from the best feature for each age group look like?

```{r}
require(ggpubr)

plot_model_preds <- function(feat_name, fits, hum_data, age="adults", mdat) {
  beta = 1 #subset(fits, AgeGroup==age & feature==feat_name)$beta
  md <- get_model_target_choice_props(get_feature_df(mdat, feat_name), hum_data, beta)
  md %>% ggplot(aes(x=chose_target, y=Model_targ_prop, color=relation)) +
    geom_abline(slope=1, intercept=0, linetype="dashed") +
    geom_point(alpha=.5) + geom_smooth(method='lm') + 
    xlab("Human Choice Probability") + 
    ylab("Model Choice Probability") +
    ggtitle(paste0(feat_name,": ",age)) + theme_minimal() + 
    xlim(0,1) + ylim(0,1)
}

ad_supp <- plot_model_preds("support_probability", fits, adult_trial_agg, age="adults", mdat)
ch_supp <- plot_model_preds("support_probability", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("support_combo", fits, adult_trial_agg, age="adults", mdat_eng)
ch <- plot_model_preds("support_combo", fits, child_trial_agg, age="children", mdat_eng)

@@

```{r}
ad <- plot_model_preds("support_std", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_std", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("support_response_sharpness_C=0.0001", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=0.0001", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("len_inverse_sharpe_ratio", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("len_inverse_sharpe_ratio", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("avg_final_radius_objects=target", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("avg_final_radius_objects=target", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("support_response_linearity_r", fits, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_linearity_r", fits, child_trial_agg, age="children", mdat)

@@

```{r}
ad <- plot_model_preds("support_sharpness_mean", fits, adult_trial_agg, age="adults", mdat_eng)
ch <- plot_model_preds("support_sharpness_mean", fits, child_trial_agg, age="children", mdat_eng)

>>>>

## Preprocessing

```{r, load-data, echo=F, FIT_MODELS=F}

@@

mdat <- mdat %>% left_join(mdat_eng)
```

# Load human data

Define helper functions

```{r, echo=F}
# selects given feature and returns wide df that can be compared with behavioral data
get_feature_df <- function(mdat, fname) {
  fdf <- mdat[,c(fname,"drop","target","relation")] %>%
    pivot_wider(id_cols = c("drop","relation"), names_from = "target", values_from = fname) %>%
    arrange(relation,drop)
}

load(file=here("paper/processed_data.RData"))

@@

evalFit <- function(feat_df, hum_df, beta, cor_objective=F, relation=NA) {
  hum_df <- get_model_target_choice_props(feat_df, hum_df, beta)
  if(!is.na(relation)) {
    hum_df <- hum_df[which(hum_df$relation==relation),]
  }
  r = with(hum_df, cor(chose_target, Model_targ_prop))
  mse = with(hum_df, sum((chose_target - Model_targ_prop)^2)) / nrow(hum_df) 
  if(cor_objective) mse = mse - r # also try to minimize -correlation
  return(mse) # return negative cor since DEoptim minimizes
}

@@

```{r fit-models-to-all-trials-mse, echo=F, eval=FIT_MODELS}
require(DEoptim)

all_trials_mse <- tibble()

@@

Could also fit separately to contain and support trials with MSE - r objective...

@@

## Fits to all trials (MSE - r)

@@

```{r, echo=F}
fit_tab <- supp_trials_mse %>% pivot_wider(names_from = Age, values_from = c(beta, mse)) %>%
  arrange(mse_adults)

@@

## Fits to containment trials (MSE)

@@

What do the predicted choice proportions from the best features for each age group look like?
Let's look at the best feature for each each group from each fit method.

```{r, echo=F, message=F}
require(ggpubr)

plot_model_preds <- function(feat_name, fits, hum_data, age="adults", mdat) {
  this_dat = subset(fits, Age==age & feature==feat_name)
  md <- get_model_target_choice_props(get_feature_df(mdat, feat_name), hum_data, beta)
  md %>% ggplot(aes(x=chose_target, y=Model_targ_prop, color=relation)) +
    geom_abline(slope=1, intercept=0, linetype="dashed") +
    geom_point(alpha=.5) + geom_smooth(method='lm', formula=y ~ x) + 
    xlab("Human Choice Probability") + 
    ylab("Model Choice Probability") +
    ggtitle(paste0(feat_name,": ",age)) + theme_minimal() + 
    xlim(0,1) + ylim(0,1) +
    annotate("text", x=.5, y=0.05, 
             label = paste0("beta = ",round(this_dat$beta,2),", MSE = ",round(this_dat$mse, 3))) # 
}
```

## Plots of other best-fitting features for all trials (MSE fits)

```{r, echo=F, message=F}
ad_supp <- plot_model_preds("support_probability", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch_supp <- plot_model_preds("support_probability", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_combo", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_combo", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r, echo=F, message=F}
ad <- plot_model_preds("support_std", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_std", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_response_sharpness_C=0.01", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_sharpness_C=0.01", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("len_inverse_sharpe_ratio", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("len_inverse_sharpe_ratio", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("avg_final_radius_objects=target", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("avg_final_radius_objects=target", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_response_linearity_r", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_response_linearity_r", all_trials_mse, child_trial_agg, age="children", mdat)

@@

```{r,  echo=F, message=F}
ad <- plot_model_preds("support_sharpness_mean", all_trials_mse, adult_trial_agg, age="adults", mdat)
ch <- plot_model_preds("support_sharpness_mean", all_trials_mse, child_trial_agg, age="children", mdat)
